### 2019-02-22 latex/chapters/intro.tex
Start adding citations

% !TEX root = ../agglo_clust_review.tex

\section{Introduction}

\begin{itemize}
\item \textbf{General intro about CNN:} big success in pixel-level image understanding tasks: boundary detection \cite{arbelaez2011contour,xie2015holistically,maninis2018convolutional}, semantic segmentation \cite{long2015fully,chen2018deeplab,kong2018recurrent}, optical flow \cite{weinzaepfel2013deepflow,dosovitskiy2015flownet}, and pose estimation \cite{wei2016convolutional,cao2017realtime}
% \SOURCE{\cite{kong2018recurrent}}
\item Many recent successful methods generate \textbf{region proposals} and classify the objects in the bounding box \cite{yang2012layered,ladicky2010and,hariharan2014simultaneous,chen2015multi,dai2016instance,liang2016reversible,he2017mask}
\item \emph{Why we do not like them}:
\begin{itemize}
\item rely on the object detector and non-maximum suppression heuristics to accurately “count” nb. of instances, 
\item difficult to train in an end-to-end manner: interface between instance segmentation and detection is non-differentiable
\item underperform in cluttered scenes (assignment carried out independently for each detection) 
\item not usable for wiry or articulated segments/objects 
\item nothing in the architecture preventing a pixel to be shared between two instances
\item number of instances is limited by nb. proposals processed by CNN (usually hundreds). 
% \item (architecture is complex and hard to tune and “debug”...?)
\end{itemize}

\item Among successful \textbf{proposal-free methods} there are: predict pixel embedding vectors \cite{kong2018recurrent,fathi2017semantic,newell2017associative,de2017semantic}; predict short- and long-range affinities \cite{liu2018affinity,wolf2018mutex,xie2015holistically}

\item In both approaches we can obtain the final segmentation by using a \textbf{graph clustering algorithm}
\item Recent work (\cite{wolf2018mutex}, more?) shows that it is better to use \textbf{repulsion and attraction} (directly predict it with the classifier, no need to define seeds or to fix a threshold given an hierarchy of clusters)
\item solving correlation clustering problem (multicut) is too expensive for this application (even using recently proposed heuristics)
\item our contributions:
\begin{itemize}
\item we propose a unified and simple formalization of Agglomerative Clustering and show how many of the recently proposed methods can be seen as special cases
\item we propose several new clustering algorithms, among which one proved to be robust in our experiments (average linkage with must-not-link-edges)
\item we compare different types of agglomeration clustering on a pixel grid graph with short and long-range connections, focusing on aspects like efficieny, robustness (and energy)

\end{itemize}
\item on other datasets (connectomics) many methods are based on multi-step pipelines first predicting superpixels
\item \textit{if we get decent scores on CREMI}: we show that also on neuro-data it is worth to skip the hand-crafted superpixel step and compute the final segmentation directly from the CNN predictions (MWS already showed it on ISBI)

\end{itemize}