### 2019-03-14 latex/chapters/4_long_range_graph_def.tex

% !TEX root = ../agglo_clust_review.tex

\section{Long-range pixel grid graph}
\begin{itemize}
\item Define predictions of the classifier (pseudo probabilities) for edge weights
\item Define possible mappings to signed costs
\item Define long-range grid graph
\item Introduce "Poisson" random graph with a given probability of long-range connections
\item Comment about the short- and long-range distinction in the MWS paper: it does not work that well in practice (applying post-processing to remove small segments is easier and better solution)
\end{itemize}