### 2020-02-17 chapters/6_conclusions.tex

% !TEX root = ../agglo_clust_review.tex
\section{Conclusion}

We have presented a novel unifying framework for agglomerative clustering of graphs with both positive and negative edge weights; and we have shown that several existing clustering algorithms, e.g. the Mutex Watershed, can be reformulated as special cases of one underlying agglomerative algorithm. This framework also allowed us to introduce new algorithms, one of which, based on an \emph{Average} linkage criterion, outperformed all the others: it proved to be a simple and remarkably robust approach to process short- and long-range predictions of a CNN applied to an instance segmentation task.
On biological images, this simple average agglomeration algorithm can represent a valuable choice for a user who is not willing to spend much time tuning complex task-dependent pipelines based on superpixels.  
% In future work, we plan to explore common theoretical properties of the algorithms included in the framework.