### 2020-11-06 extra/submissions_uploads/arXiv_submission/chapters/6_conclusions.tex

% !TEX root = ../ms.tex
\section{Conclusion}
We have presented a novel unifying framework for agglomerative clustering of graphs with both positive and negative edge weights and we have shown that several existing clustering algorithms, e.g. the Mutex Watershed \cite{wolf2018mutex}, can be reformulated as special cases of one underlying agglomerative algorithm. This framework also allowed us to introduce new algorithms, one of which, based on an \emph{Average} linkage criterion, outperformed all the others: It proved to be a simple and remarkably robust approach to process short- and long-range predictions of a CNN applied to an instance segmentation task.
On biological images, this simple average agglomeration algorithm can represent a valuable choice for user who is not willing to spend much time tuning complex task-dependent pipelines based on superpixels.  
In future work we plan to extend the comparison to other types of non-image graphs and explore common theoretical properties of the algorithms included in the framework.