### 2021-03-04 chapters/6_conclusions.tex

% !TEX root = ../agglo_clust_review.tex
\section{Conclusion}
We have presented an unifying framework for agglomerative clustering of graphs with both positive and negative edge weights; and we have shown that some existing clustering algorithms, e.g. the Mutex Watershed, can be reformulated as special cases of this framework. 
We have analyzed several theoretical and empirical properties of the algorithms included in this framework. On instance segmentation, algorithms based on an average linkage criterion outperformed all the others: they proved to be simple and remarkably robust approaches to process short- and long-range predictions of a CNN.
 % applied to an instance segmentation task.
On biological images, these simple average agglomeration algorithms achieve state-of-the-art results without requiring the user to spend much time tuning complex task-dependent pipelines based on super-pixels.
% can represent a valuable choice for a user who is not willing to spend much time tuning complex task-dependent pipelines based on super-pixels.  
% In future work, we plan to explore common theoretical properties of the algorithms included in the framework.
% On biological images, this simple average agglomeration algorithm superpixels