### 2021-03-10 chapters/0_abstract.tex
Rewrite supplementary material and add proofs

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a new simple theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations and introduce new algorithms for combinations that have not been studied before. 
We study both theoretical and empirical properties of these combinations and prove which ones of them define an ultrametric on the graph.
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
Lastly, we find that some of the algorithms included in our framework, when combined with the predictions from a CNN model, result in a simple instance segmentation pipeline that achieves state-of-the-art performances on the competitive CREMI 2016 EM segmentation benchmark.

% On stochastic block model problems, GASP compares favorably to spectral clustering. 
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}