### 2021-03-17 chapters/6_conclusions.tex
Add comments from Anna

% !TEX root = ../agglo_clust_review.tex
\section{Conclusion}
We have presented a unifying framework for agglomerative clustering of graphs with both positive and negative edge weights. This framework allowed us to explore new combinations of constraints and linkage criteria and to perform a consistent evaluation of all algorithms in it. 
% and we have shown that some existing clustering algorithms, e.g. the Mutex Watershed, can be reformulated as special cases of this framework. 
We have then analyzed several theoretical and empirical properties of these algorithms. On instance segmentation, algorithms based on an average linkage criterion outperformed all the others: they proved to be simple and robust approaches to process short- and long-range predictions of a CNN.
 % applied to an instance segmentation task.
On biological images, these simple average agglomeration algorithms achieve state-of-the-art results without requiring the user to spend much time tuning complex task-dependent pipelines based on super-pixels.
% can represent a valuable choice for a user who is not willing to spend much time tuning complex task-dependent pipelines based on super-pixels.  
% In future work, we plan to explore common theoretical properties of the algorithms included in the framework.
% On biological images, this simple average agglomeration algorithm superpixels